A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-04-07 faculty.tex
updates

\chapter{Tenure Package and Application}

@@

\subsection{Letters}

>>>>

\chapter{Tenure Package and Application}
After 5 or 6 years, it is time for you to go up (or apply) for tenure.  You will need to prepare a tenure package and submit it to your department.  The department will then review your package and vote on whether to recommend you for tenure.  If the department votes in your favor, the package will then go to the college, and then to the university.  The university will then make the final decision on whether to grant you tenure.

This document shows my experience in preparing my tenure package and applying for tenure. I will share my timeline, the package (statements, letter writer sections), and the process.  I will also share some advice on this process.

\section{Timeline}
Typically you will first mention to your department chair that you are planning to go up for tenure. The department chair then assign you a mentor who will help you with the process. You will apply for tenure in the Fall semester, and the decision will be made in the Spring semester.

My own timeline is as follows:

\begin{itemize}
    \item Fall 2021: Moved to GMU (I started at UNL in 2016)
    \item (late) Fall 2022: I told my department chair that I plan to go up for tenure in 2023.
    \item Spring 2023: I started preparing my tenure package, which includes writing statements and think reference letters.  I also took a semester off because GMU has a pre-tenure study break policy.
    \item Summer 2023: I submitted my tenure package.
    \item Fall 2024: I gave a tenure talk. Department voted in my favor, and then the college voted in my favor, and the decision goes up to the university.
    \item Spring 2024: In March, I got the official notification that I got tenure and promoted to Associate Prof., beginning in Fall 2024.
\end{itemize}

@@

\subsection{Letters}
This is a huge part of the tenure package. You will need to ask for letters from senior people in your area.  Someone who can vouch for your research, teaching, and services.

You will need 4-6 letters, however you will not be the one asking for these letters, but your department will. However you can suggest who to ask. Typically, there will be two list of letter writers: one list consists of people suggested by you and the other list has people suggested by the department (e.g., people who work in your field in your dept).

Originally I asked several senior people in my area to write letters for me. Then I found out that there is a rule at GMU that I cannot ask people whom have collaboration with me in the last 7 years!!! That essentially make most of my letter writers invalid.  Moreover, GMU (and likely most other universities) prefer letter writers not from my list (because people on my list are likely to be biased), and has a rule that at least 60\% of the letter writers must be from the department's list. Also, there's a good chance that a letter writer will decline the invitation, so there should be more than 4-6 names in the lists.

After talking to various people in my dept and outside, I decide to put just 2 names in my list and let the department decides the rest. I even thought about leaving my list empty but then that seems to be too weird and I don't want to give the impression that I don't have any friends in the world.  Essentially this means that likely none or 1 of

The reason is because I feel I don't have too many "enemies" out there that would write bad things about me.  Those that don't know me well will likely just decline the letter invitation.  Turned out to be a good decision because later I learn that all my letters wrote very positive things about me.  I will never know who wrote the letters, but I am very grateful to them.
